---
title: "ichimoku: Cloud Charts - The Reference Manual"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{ichimoku: Cloud Charts - The Reference Manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height = 5
)
```

## Ichimoku Kinko Hyo

### Definition

The 一目均衡表 **Ichimoku Kinko Hyo** [cloud chart] consists of the following chart lines added to a candlestick chart:

1. 転換線 **Tenkan-sen** [conversion line]: the mid-point of the highest high and lowest low for the past 9 periods *(including the current period)*

2. 基準線 **Kijun-sen** [base line]: the mid-point of the highest high and lowest low for the past 26 periods *(including the current period)*

3. 先行スパン 1 **Senkou span A** [leading span A]: the mid-point of Tenkan-sen and Kijun-sen plotted ahead 26 periods *(including the current period)*

4. 先行スパン 2 **Senkou span B** [leading span B]: the mid-point of the highest high and lowest low for the past 52 periods *(including the current period)*, plotted ahead 26 periods *(including the current period)*

5. 遅行スパン **Chikou span** [lagging span]: the current period closing price plotted behind 26 periods *(including the current period)*

The 雲 **kumo** [cloud] is the area bounded by Senkou span A and Senkou span B (usually shaded on a chart).

### Interpretation

Ichimoku Kinko Hyo can be translated as 'one-glance equilibrium chart'. It is designed to allow the price action and market structure of financial securities to be determined visually 'at-a-glance'.

For example in a strongly upwards-trending market, the candlesticks would be above the Tenkan-sen, which would be above the Kijun-sen, which in turn would be above the cloud, and the Chikou span may not have anything above it.

The lines and the cloud represent dynamic support and resistance zones relative to the price candles. Generally, the thicker the cloud, the tougher the support/resistance. In our previous example, if the price now reverts downwards, it can expect support first at the Kijun-sen, then the Tenkan-sen and finally the cloud itself.

When the price enters the cloud, i.e. it is between the cloud base and cloud top, it is an important area to gauge whether there is sufficient momentum for the price to break through the cloud or whether the cloud eventually provokes a reversal.

More subtle interpretations involve the Chikou span in particular and its action in relation to the cloud lines as well as the candles.

It is outside the scope of this vignette to provide a full tutorial on the use of Ichimoku Kinko Hyo.

### Context

Ichimoku analysis is the latest refinement in candlestick charting techniques, which also originated from Japan back in the 18th century. Actually developed during the mid-20th century, it gained popularity especially from the late 1990s onward, and is now used on trading floors worldwide.

The ichimoku time periods have traditionally been calculated as 9, 26 and 52 based on manual data analysis performed in Japan in a pre-computer age where there was a 6-day working week resulting in 26 average trading days in a month. Although this bears little relevance to the current day, the use of these time periods has persisted as an 'industry norm' or 'accepted practice'. To use other periods would be meaningless in a sense as everyone uses these parameters and 'market psychology' can and often does create its own realities, independent of any fundamentals.

However, there is no reason for the technique not to evolve, and to reflect changing trading realities perhaps other parameters will become more relevant in the collective psychology. For this reason, the length of these periods can be freely set in the 'ichimoku' package. However please do so only with a strong sense of caution and note that using other periods invalidates the traditional interpretations of Ichimoku Kinko Hyo.

Finally, the use originated with daily candlesticks, and the most valid interpretation remains for daily data. However, it is equally used today for both shorter intra-day, e.g. 4-hour or hourly, and longer, e.g. weekly or monthly, charts.

## A Typical Workflow
```{r setup}
library(ichimoku)
```

### Data

ichimoku works with data frames and tabular data, and interfaces with other R packages providing financial price data. The `ichimoku()` function is able to handle as input xts, data.frame, data.table, tibble, matrix, and other data.frame compatible objects. For further details see the section [compatible formats](#compatible-formats).

```{r loaddatanotrun, eval = FALSE}
# ichimoku is designed to interface well with other packages, for example:
cloud <- ichimoku(quantmod::getSymbols("C"))
```

ichimoku also incorporates an interface with the OANDA fxTrade API to return data for major currencies, metals, commodities, government bonds and stock indices. Please refer to the [OANDA vignette](xoanda.html).

```{r oandadatanotrun, eval = FALSE}
# ichimoku can create clouds directly from OANDA data, for example:
cloud <- ichimoku(oanda("USD_JPY"))
```

The requirement on the input data is minimal: a series of prices indexed by valid date-times. 

```{r viewdata}
# Simulated OHLC price data is assigned to data frame 'TKR':
TKR <- sample_ohlc_data
head(TKR)
```

**Date-time index**

- Should be in a format convertible to POSIXct

- ichimoku recognises indexes in columns with names containing 'date', 'time' or 'index', or else the row names of the dataset

**Open/High/Low/Close (OHLC) data** should always be used where available.

- ichimoku recognises data in columns with names containing 'open', 'high', 'low' and 'close'. These do not have to be nicely-formatted as in the example above and could also contain other text such as the ticker name. For further details see the section [data validation](#data-validation)

- If only HLC price data is available, ichimoku will take the opening price to be the closing price of the previous period. *This means that the candlesticks will not be accurate, but calculation of the cloud lines is not affected*

**Single price data** can also be used for producing an *approximate* cloud chart if, for example, OHLC data is not available or because the price is a calculated price such as VWAP.

- If ichimoku does not find HLC data in the dataset, it will proceed to seek for single price data in columns with names containing 'price', 'value', or 'close'

- ichimoku will construct a pseudo-OHLC series from the single price. It does this by assigning the opening price to be the closing price of the previous period, and the high and low to be the max and min respectively of the opening and closing prices

- As the constructed OHLC series does not contain real high/low values, *the ichimoku cloud charts calculated in this way are not accurate and serve as approximations only*

**Note on data retrieval:** it is optimal to retrieve data starting from slightly further back than the period being analysed. This is as the cloud lines require several values to be calculated and hence there will be missing values at the start of the data. Run `ichimoku()` on the full dataset and then subset the data window when plotting to ensure that there is a full cloud for the entire chart.

### ichimoku()

`ichimoku()` is the main function of the ichimoku package. It takes the input data object and returns an ichimoku object, which inherits the classes 'ichimoku', 'xts' and 'zoo'.

Optional arguments:

- `ticker` a ticker to identify the instrument, otherwise this is set to the name of the input object
- `periods` [default c(9, 26, 52)] a vector defining the length of periods used for the cloud. This parameter should not normally be modified as using other values would be invalid in the context of traditional Ichimoku analysis
- `...` additional arguments, for instance 'holidays', passed along to `tradingDays()` for calculating the future cloud on daily data

An ichimoku object is returned by calling `ichimoku()` on the data:

```{r ichimoku}
cloud <- ichimoku(TKR)

print(cloud[100:110,], plot = FALSE, digits = 4)
```

A section (rows 100 to 110) of the ichimoku object is shown above. Note that print is called setting 'plot = FALSE' to return the data only. ichimoku objects employ a custom S3 print method which by default plots the cloud chart in the graphical device as well as returning the data to the console.

```{r ichimokunotrun, eval = FALSE}
# to view chart as well as data, simply issue:
cloud
```

#### The ichimoku object specification

- Index:
  + `index(object)` - date-time index [POSIXct]

- Columns [numeric]:
  + `object$open` - opening price
  + `$high` - high price
  + `$low` - low price
  + `$close` - closing price
  + `$cd` - candle direction (-1 = down, 0 = flat, 1 = up)
  + `$tenkan` - Tenkan-sen
  + `$kijun` - Kijun-sen
  + `$senkouA` - Senkou span A
  + `$senkouB` - Senkou span B
  + `$chikou` - Chikou span
  + `$cloudT` - cloud Top (max of senkouA, senkouB)
  + `$cloudB` - cloud Base (min of senkouA, senkouB)

- Attributes:
  + `attributes(object)$periods` - parameters used to calculate the cloud [integer vector of length 3]
  + `$periodicity` - periodicity of the data in seconds [numeric]
  + `$ticker` - instrument identifier [character]
  
The ticker is set by default to the input object name or can be overridden by specifying the 'ticker' argument to `ichimoku()`.

#### ichimoku.ichimoku()

`ichimoku()` may be used on an ichimoku object to re-calculate the cloud values using the price data contained within. If 'ticker' is specified as an argument, this will overwrite the ticker stored in the original ichimoku object.

Calling `ichimoku()` on an ichimoku object containing a strategy will return the ichimoku object to its original state without the strategy. See the [strategies vignette](strategies.html) for information on strategies.

#### Working with ichimoku objects

As an ichimoku object inherits the 'xts' and 'zoo' classes, all the existing methods for these classes should work seamlessly. For convenience, the following functions are re-exported by ichimoku:

*from 'zoo':*

- `index()` to extract the index of an ichimoku object
- `coredata()` to extract the columns of an ichimoku object as a numeric matrix (without the date index)

*from 'xts':* 

- `xts()` to create an 'xts' object from data and a date-time index, use `xts(data, index)`

If modifications have been made, call `ichimoku()` on the modified object to re-create the ichimoku object according to the correct specification.

For additional functionality, you may consider loading the 'xts' package to access a wide range of methods available for 'xts' objects, for example changing the periodicity of the data.

#### Defining custom holidays

When calculating the future cloud for daily data, `ichimoku()` makes a call to the `tradingDays()` helper function to ensure that future dates fall on trading days (i.e. not weekends or holidays). By default, New Year's and Christmas day are defined as holidays as these are for all purposes universal across markets. These defaults should be acceptable for general use.

If you wish to define other dates as non-trading days, additional custom holidays may be supplied via a 'holidays' argument to `ichimoku()` which is then passed on to the `tradingDays()` function.

If the relevant market trades on a 24/7 basis, the parameter `noholidays` can be specified. This will bypass the logic of `tradingDays()` and use all dates for calculating the future cloud.

```{r holidays, eval = FALSE}
# Holidays can be specified directly via a vector of dates:
ichimoku(TKR, holidays = c("2020-01-13", "2020-02-11", "2020-02-24"))

# Or via a functions that returns a vector of dates (e.g. from the 'timeDate' package):
ichimoku(TKR, holidays = timeDate::holidayLONDON())
ichimoku(TKR, holidays = timeDate::holidayNYSE())

# For a market that trades 24/7:
ichimoku(TKR, noholidays = TRUE)

```

### plot()

ichimoku offers the choice of 2 visualization systems: static (default) and interactive (reactive charts using R Shiny).

#### 1. Static Plots

The default plot function produces static plots, which have the advantage of being easily exportable to pdf or image formats in high resolution.

ichimoku implements a custom S3 method for ggplot2's `autoplot()` function for objects of class 'ichimoku'. Although autoplot can be called directly on an ichimoku object, it is more convenient to call `plot()` instead, which prints the output of autoplot.

```{r plot}
plot(cloud)
```

As the implementation is via ggplot2, the user is able to further amend or customise the resulting plot using any ggplot2 methods. The most straightforward method would be to assign the output of `plot()` to a 'ggplot' object, after which further elements or layers can be added or modified using ggplot2's `+` operator.

```{r plotnotrun, eval = FALSE}
# Example code for incrementally updating ggplot2 layers:
plot <- plot(cloud)
plot + ggplot2::theme_classic()
```

The 'ggplot' object can be further used to produce an interactive chart that can be embedded in HTML output created by R Markdown, by using the 'plotly' package as per the below:

```{r iplotnotrun, eval = FALSE}
# For a plotly interactive plot:
plotly::ggplotly(plot)
```

#### 2. Interactive Plots

Interactive plots require the 'shiny' package to be installed. If Shiny is not available, ichimoku will issue an informational message, but not install the package automatically as it is recognised that Shiny may not be available in all environments.

To produce an interactive plot, use ichimoku's `iplot()` function. An R Shiny app will be launched, allowing full further customisation of the chart interactively, including selecting the time window etc.

In a live analysis environment, as opposed to creating charts for reports, it can be preferable to work with interactive plots. This is due to the ease of being able to customise the chart coupled with easy access to the data provided by the cursor infotip.

- hover over the chart to read off data values at any given date-time
- cursor guides provide easy calibration of date-times and price levels
- select periods of interest on the timeline to subset / enlarge

#### Plot customisation

A plot with default parameters is shown above, but the plot functions contain built-in customisations to allow for most envisioned use cases. The example below demonstrates some of the arguments that can be supplied to customise the plot.

```{r plot2}
plot(cloud, window = "2020-05-01/2020-12-01", ticker = "TKR Co.", message = "一目均衡表", theme = "dark")
```

The following arguments for customisation are shared across `plot()` and `iplot()`:

- `window` (optional) indispensable in a proper workflow. `ichimoku()` is run on a larger dataset than that of interest and at the visualization stage the data is subset to the window of interest. Supply as an ISO-8601 compatible range string in the format used for 'xts' objects, for example the range '2020-02-15/2020-08-15' or '2020-02-15/' (from 15 Feb 2020) or '/2020-08' (until end-Aug 2020) or simply '2020' to select all of the year 2020
- `ticker` (optional) used to supply a different ticker to that stored in the ichimoku object, or alternatively a longer-form name or other text that will appear in the chart heading
- `message` (optional) used to specify a chart message to display under the title
- `theme` [default 'original'] the following themes are also available: 'dark', 'solarized' and 'mono'
- `strat` [default TRUE] if true, the periods for which a strategy results in a market position will be shaded (if the ichimoku object contains a strategy) with the strategy printed as the chart message (if a message is not already specified). See the [strategies vignette](strategies.html) for more information on working with strategies.

`plot()` takes additional parameters as follows:

- `...` additional arguments passed along to the print method for 'ggplot' objects.

`iplot()` takes the following additional parameters:

- `...` additional parameters passed along to the 'options' argument of `shiny::shinyApp()`.
- `launch.browser` [default TRUE] If TRUE, the system's default web browser will be launched automatically after the app is started. The value of this argument can also be a function to call with the application's URL. To use the default Shiny viewer in RStudio, please specify
`getOption("shiny.launch.browser")`.

### archive()

After completing an analysis, the resulting objects can be stored directly to disk. The function `archive()` combines R's native serialisation capability with cryptographic verification of data integrity.

Archive files created by this function are stored in the native RData format for maximum portability and reliability. They may be read using the base R `load()` function on machines without 'ichimoku' installed.

#### Writing

`archive()` for write operations will take 2 arguments: the object to be written, and the file path/name of the file to be stored. A confirmation is printed to the console that the file has been written.

#### Reading

`archive()` for read operations takes a single argument: the file path/name of the file to be read. The return value of the function can be assigned to an object. A confirmation is printed to the console once the file has been read.

#### Data Integrity Verification

Data integrity verification is performed by the cryptographic-grade sha256 hashing algorithm from the 'openssl' package. If 'openssl' is not available, the archive function still works in all respects but without the authentication capability.

When an archive is written, the serialised object is hashed and the hash is also stored in the archive. The sha256 hash value is printed to the console as confirmation.

When an archive is read back, the sha256 hash of the restored object is checked against the hash of the original stored in the archive. If identical, a 'data verified' message is printed to the console along with the authenticated sha256 hash.

## Troubleshooting

### Compatible formats

`ichimoku()` is a generic function that provides S3 methods for objects of class 'xts', 'data.frame' and 'matrix'.

Other formats such as 'data.table' and 'tibble' that inherit the class 'data.frame' should also work with `ichimoku()`.

ichimoku has been designed to be fully pipeable and interfaces well with other R packages that return financial data. As an example, the `getSymbols()` function from the 'quantmod' package currently exhibits non-standard R behaviour which assigns the data into an object but returns a text string containing the name of the object. ichimoku is able to handle this automatically such that a chart can be created without the need to save intermediate objects:

```{r pipe, eval = FALSE}
# Using R 4.1's new pipe operator:
quantmod::getSymbols("C") |> ichimoku() |> plot()
# Or equally using the 'magrittr' pipe:
quantmod::getSymbols("C") %>% ichimoku() %>% plot()
```

### Data validation

ichimoku performs the following data validation steps before attempting to compute the cloud values. Failure at any step causes `ichimoku()` to halt.

1. Searches column names of the data object for the first column containing 'index', 'date' or 'time' (not case-dependent). If such a column is found but the data is not convertible into POSIXct format, an error is returned

  + *If not found:* searches in row names of the data object if present. If the row names do not contain data that is convertible into POSIXct format, an error is returned. Note: 'xts' objects are always indexed by a valid timestamp and ichimoku handles this automatically

2. Searches for HLC price data in columns with names containing 'high', 'low' and 'close' (not case-dependent) respectively

  + *If found:* Searches for opening price data in a column with name containing 'open' (not case-dependent) and *if not found* sets the opening prices to be the closing prices of the previous period

  + *If not found:* Searches for a single price series in columns with names containing 'price', 'value' and 'close' (not case-dependent), for which a pseudo-OHLC series will be reconstructed

3. Checks that the dataset is longer than the medium cloud period. If this is not the case then none of the cloud lines can be calculated

There may however still be other issues with the input data that prevent cloud values from being computed, not handled explicitly by `ichimoku()`. For example, if:

- The date-time index is in a headerless column, or in a column where the name does not contain 'index', 'date' or 'time' (not case-dependent). Please rename the column to 'index' in such cases

- The correct date-time index is in the row names, or index of an 'xts' object, but there are also columns that contains the word 'index', 'date' or 'time'. Please rename the columns that do not serve as the date-time index so that the correct date-time index can be detected

- There is more than one column that contains the word 'index', 'date' or 'time'. Please either rename the columns that do not serve as the date-time index, or move the correct date-time column ahead of the other columns so that it is picked up first

- HLC price data is contained in headerless columns or in columns labelled, for example, 'Op', 'Hi', 'Lo', 'Cl'. Please rename the columns 'open', 'high', 'low' and 'close' in such cases

- There are NA values within the price data. This will affect the algorithm calculating the cloud lines in the window immediately following such NA values. Please check the data and manually correct NA values before invoking `ichimoku()`

### API modifications

#### Since version 0.x

- The first ichimoku object specification was released in version 0.3.51 - ichimoku objects created in versions prior to 0.3.51 will no longer work correctly with newer versions of the package
- Upgrade to the latest package version and run `ichimoku()` on previously-created objects to re-create them according to the new specification (data is preserved)

## Dependencies

The dependency on other packages is designed to be minimised to the extent possible. ichimoku currently has the following external package dependencies:

- curl - for interfacing with external APIs
- ggplot2 - the widely-used R graphics system for visualization
- jsonlite - for parsing JSON data retrieved from external APIs
- rlang - R language requirement arising from the use of 'ggplot2'
- xts - ichimoku objects are based on the 'xts' format for time series
- zoo - 'xts' is built on 'zoo' technology

Compile time:

- cpp11 - provides the C++ headers for various internal functions used by ichimoku

Optional (provides additional features):

- bslib - for full theming of 'shiny' applications
- keyring - for saving API keys in the system credential store
- openssl - for verification of the data integrity of archived objects
- shiny - for interactive visualizations

## References

Sasaki, H. 佐々木 英信 (1996), *一目均衡表の研究 [ichimoku kinkouhyou no kenkyuu]*. Tokyo, Japan: Toushi Radar.

Gao, C. (2021), *ichimoku: Visualization and Tools for Ichimoku Kinko Hyo Strategies*. R package version 1.1.1, <https://CRAN.R-project.org/package=ichimoku>.

---
